---
title: |
  | Final Project
  | DS 807: Modeling Unstructured Data
author: |
  | Your Name Goes Here
output: html_document
---

## Data Requirements:

- You can pick any type of data that satisfies **at least one** of the following criteria:

    1. Text Data
    2. Image Data
    3. Unsupervised Data


- Some sources are:

    - Kaggle <https://www.kaggle.com/datasets>
    - UCI Machine Learning Repository <https://archive.ics.uci.edu/ml/index.php>
    
- Read your data in R.

```{r}
library(tidytext)
library(dplyr)
library(ISLR) 
library(factoextra)
library(cluster)
library(tidyverse)
library(tidyr)
library(RColorBrewer)
library(EnvStats)
library(ggpubr)
```


```{r}
spotify<-read.csv("C:/Masters/Unh/Unstruc_data/Project/data/dataset_Year.csv", header = TRUE, sep = ",")

sum(is.na(spotify))
spotify=drop_na(spotify)
sum(is.na(spotify))
```

## The grading rubric can be found below:

+----------------+---------------+--------------------+-----------------------+
|                | R code        | Decision/Why       | Communication         |
|                |               |                    |  of findings          |
+================+===============+====================+=======================+
| Percentage of  | 30%           | 35%                | 35%                   |
| Assigned Points|               |                    |                       |
+----------------+---------------+--------------------+-----------------------+


- **Decision/why?**: Explain your reasoning behind your choice of the procedure, set of variables and such for the question. 

    - Explain why you use the procedure/model/variables
    - To exceed this criterion, describe steps taken to implement the procedure in a non technical way.


- **Communication of your findings**: Explain your results.

    - Explain why you think one model is better than the other.
    - To exceed this criterion, explain your model in a non technical way.

## Note

- Since there is a great range of potential data, the instructions are written in a general way. If some steps do not make sense in your case, please reach out to verify.

## Part 1: Exploratory Data Analysis (20 points)

1. Explain the purpose of the analysis *for your data*, i.e., what are you trying to achieve?

2. Check for existence of NA's (missing data)

3. Use appropriate plots for EDA, i.e., word counts for text data.

4. Do you need to scale your data or do you need dimension reduction? If so, perform a principle components analysis on a scaled, or not-scaled data depending on your needs.

```{r}
#Top 10 genres by Popularity
spotify%>%
  select(popularity,track_genre)%>%
  group_by(track_genre)%>%
  summarize(Avg_Popularity=mean(popularity))%>%
  arrange(desc(Avg_Popularity))%>%
  top_n(10, Avg_Popularity)%>%
    ungroup()%>%
  ggplot(aes(y = reorder(track_genre, Avg_Popularity), x = Avg_Popularity)) +
  geom_bar(aes(fill = Avg_Popularity), stat = "identity") +
  scale_fill_gradient(low = "#F9CCDA", high = "#5A3DDA") +
  labs(title = "Top 10 most popular genre",
       y = "genre",
       x = "Popularity") +
  theme_minimal()

#Top 10 songs covering most of the genres
spotify%>%
  select(track_name,track_genre)%>%
  group_by(track_name)%>%
  summarise(Number_of_genre=n())%>%
  select(track_name,Number_of_genre)%>%
  group_by(track_name)%>%
  top_n(1,Number_of_genre)%>%
  ungroup()%>%
  arrange(desc(Number_of_genre))%>%
  top_n(10,Number_of_genre)%>%
  ggplot(aes(y = reorder(track_name, Number_of_genre), x = Number_of_genre)) +
  geom_bar(aes(fill = Number_of_genre), stat = "identity") +
  scale_fill_gradient(low = "#a9CCDA", high = "#0A3DDA") +
  labs(title = "Top 10 most popular songs",
       y = "Song",
       x = "Popularity") +
  theme_minimal()

#Top 2 Songs in top 10 genres
top_10_genre=spotify%>%
  select(popularity,track_genre)%>%
  group_by(track_genre)%>%
  summarize(Avg_Popularity=mean(popularity))%>%
  top_n(10,Avg_Popularity)

spotify%>%
  select(popularity,track_name,track_genre)%>%
  filter(track_genre == top_10_genre$track_genre)%>%
  group_by(track_genre)%>%
  slice_max(n=2,popularity, with_ties = FALSE)%>%
  mutate(label = paste0(track_name))%>%
  ggplot(aes(y = reorder(track_name, popularity), x = popularity)) +
  geom_bar(aes(fill=factor(track_genre)),position = position_dodge(),stat = "summary") +
  geom_label( x=20, aes(label=label), size=2) +
  labs(title = "Top 2 songs for top 10 popular genres",
       y = "songs",
       x = "Popularity") +
  theme_minimal()+
  scale_fill_brewer(palette="BrBG")+
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

#Popularity % accross whole data
spotify_temp<-spotify%>%
  mutate(Popularity = ifelse(popularity<=25,"Least Popular",ifelse(popularity>25 & popularity<=50,"Not so Popular",ifelse(popularity>50 & popularity<=75, "Popular", "Most Popular"))))

spotify_temp%>%
  group_by(Popularity)%>%
  summarise(Count=n())%>%
  mutate(Popularity=as.factor(Popularity),percentage=round(Count/sum(Count)*100,2),ymax =cumsum(percentage),ymin = c(0, head(ymax, n=-1)),LabelPosition = (ymax + ymin)/2, label = paste0(Popularity, "\n value: ", Count, "\n", percentage,"%"))%>%
           ggplot(aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Popularity)) +
  geom_rect() +
  geom_label( x=4, aes(y=LabelPosition, label=label), size=4) +
  scale_fill_brewer(palette=4) +
  coord_polar(theta="y") +
  xlim(c(2, 4)) +
  theme_void()+
  theme(legend.position = "none")

ggplot(spotify,aes(loudness,energy)) +
   geom_point(col="violet")+
  geom_smooth(model="lm",col="darkblue")+
  stat_regline_equation(label.y = 0.8, aes(label = ..eq.label..)) +
  stat_regline_equation(label.y = 0.9, aes(label = ..rr.label..))

```
## Part 2: Clustering (20 points)

1. Develop a clustering algorithm for your data: Choose from topic models, k-means, k-medoids, hierarchical, or DBSCAN. 

```{r}
#taking variables for clustering

set.seed(1994)
spotify_v3<-sample(nrow(spotify),size=nrow(spotify)*1)
spotify_v4<-spotify[spotify_v3,]
spotify_v5<-spotify_v4

clean<-spotify_v5%>%
  select(-track_genre,-ID)

clean<-distinct(clean)

spotify_knn<-clean%>%
  select(where(is.numeric),-popularity,-key,-mode,-time_signature,-energy)%>%
  scale()

head(spotify_knn)

spotify_knn<-spotify_knn%>%
  scale()
```


```{r}
#Finding clusters : ELBOW METHOD

#The Elbow method looks at the total WSS as a function of the number of clusters: One should choose a number of clusters so that adding another cluster doesn’t improve much better the total WSS.

fviz_nbclust(spotify_knn, kmeans, method = "wss",k.max=10)


#We choose k=3 after which the wss increases
```


```{r}
#K-means Clustering Method
set.seed(1234)
km=kmeans(spotify_knn, 3)
fviz_cluster(km, data =spotify_knn,palette = "jco", ggtheme = theme_minimal())

```

```{r}
#clusterplot
fviz_cluster(km, data = spotify_knn, geom = c("point"),ellipse.type = "euclid")
```

```{r}
# cluster assignment
cluster<-km$cluster
clean$cluster <- as.factor(km$cluster)

song_clustered<-clean%>%
  select(artists,track_name,cluster)
song_clustered
```


2. Explain your choices on model parameters, i.e. k, eps, minpts, and communicate your results.

We tried to find optimal number of clusters by using WSS and silhouette method. Silhouette method showed 2 clusters. From WSS method, we observed that anymore clusters from2 is good. So, we decided to take k=3.
We ran k-means clustering method on the data. For this, we removed all non-numeric variables from the data. The compactness is 27.8% by which they are similar to each other in the cluster.

Cluster 1: Danceable, loud and high energy and valance (positive, cheerful songs)  
Eg: Sucker by Jonas Brothers, Naaka Mooka, Vijay Antony

Cluster 2: showing high energy, loudness,liveness and high tempo characteristics (inspirational song) - 
Eg: Story of My Life by One Direction

Cluster 3: This cluster has songs that are high in accoustic, instrumental and comparably less duration. This type of music emphasizes simplicity in its lyrics, harmonies, melodies or melancholies. 
Eg: 


## Part 3: Mixture Models (20 points)

1. Apply a mixture model based clustering to your data.

2. Explain your choices on model parameters, and communicate your results.

```{r}
set.seed(1234)

spotify_v4<-spotify%>%
  select(-ID,-popularity,-key,-mode,-time_signature)

spotify_k<-sample(nrow(spotify),size=nrow(spotify)*0.03)
spotify_pre=spotify_v4[spotify_k,]
track_name=spotify_pre$track_name  

spotify_pre=spotify_pre%>%
  select(where(is.numeric))%>%
  scale()

mix_m=Mclust(spotify_pre)

summary(mix_m, parameters=TRUE)

cluster <- mix_m$classification

plot(mix_m, what="classification")

param_comp_1=mix_m$parameters$pro[1]
param_comp_2=mix_m$parameters$pro[2]
param_comp_3=mix_m$parameters$pro[3]
param_comp_4=mix_m$parameters$pro[4]
param_comp_5=mix_m$parameters$pro[5]
param_comp_6=mix_m$parameters$pro[6]
param_comp_7=mix_m$parameters$pro[7]
param_comp_8=mix_m$parameters$pro[8]
param_comp_9=mix_m$parameters$pro[9]

spotify_pre=as.data.frame(spotify_pre)

spotify_c<-spotify_pre %>%
  mutate(Cluster=factor(mix_m$classification),Track_name=track_name)

#spotify_c<-spotify_pre %>%
 # mutate(Track_name=track_name)

spotify_c %>%
  group_by(Cluster) %>%
  mutate(Number=row_number()) %>%
  ggplot(aes(x=Number, y=Cluster, col=Cluster)) + 
  geom_text(aes(label=Track_name), size=3)+
  theme(legend.position = "none")
```

## Part 4: Deep Learning (20 points)

1. Apply a type of neural network algorithm to your data.

```{r}
# Load required packages
library(tidyverse)
library(caret)
library(keras)
```


```{r}
# Load Spotify dataset
# Define input and output variables
clean_2<-spotify%>%
  select(-track_genre,-ID)

clean_2<-distinct(clean_2)

# Split data into training and validation sets
spotify_ffnn<-clean_2%>%
  select(-c("track_id","popularity","artists","album_name","track_name","explicit","Year"))
```


```{r}
set.seed(1994)
index = sample(1: nrow(spotify_ffnn), round(0.7*nrow(spotify_ffnn)))
train = spotify_ffnn[index,]
test = spotify_ffnn[-index,]

nrow(spotify_ffnn)
nrow(train)
nrow(test)


x_train <- train[,1:4]
y_train <- as.matrix(train[,5:13])

x_test <- test[,1:4]
y_test <- as.matrix(test[,5:13])


#max minus min scale
maxs=apply(x_train[,1:4], 2, max)
mins=apply(x_train[,1:4], 2, min)
x_train<-scale(x_train[,1:4],center = mins, scale = maxs - mins)

maxt=apply(x_test[,1:4], 2, max)
mint=apply(x_test[,1:4], 2, min)
x_test<-scale(x_test[,1:4],center = mint, scale = maxt - mint)


```


```{r}
# Define MLP model architecture
model=keras_model_sequential()
model %>%
  layer_dense(units=64, activation='relu', input_shape=ncol(x_train)) %>%
  layer_dropout(rate = 0.4) %>%
  layer_dense(units = 32, activation = 'relu') %>%
  layer_dropout(rate = 0.4) %>%
  layer_dense(units =9, activation='linear')

summary(model)
```



```{r}
# Compile model
model %>% 
  compile(loss = 'mse',
          optimizer = 'rmsprop',
          metrics=c('accuracy', 'mae')
          )

# Train model
history<-model %>% 
  fit(x_train, y_train, epochs = 50, batch_size = 64,verbose=1,validation_split = 0.2)

model %>% 
  evaluate(x_test, y_test)

#Prediction
pred<-model %>%
  predict(x_test)
```

```{r}
temp<-as.matrix(pred)
plot(y_test[,2],col="grey",xlab = "Test Values", ylab = "Predictions", main = "Predictions vs. Test Values")
points(temp[,2], col="brown")

```

2. Explain your choices on model parameters, and communicate your results.


## Part 5: Conclusion (20 points)

1. (10 points) Based on the purpose of your analysis stated in Part 1, which analysis did a good/better/satisfactory job? How do you think you can improve the analysis?

    
2. (10 points) What are your learning outcomes for this assignment? Please focus on your learning outcomes in terms of analysis, model interpretations, and R skills - it is up to you to include this part in your presentation or not.



```{r}
km$centers
Clusters=c("Cluster-1", "Cluster-2", "Cluster-3")
Characteristics=c("Danceability, Loudness", "Liveness, Loudness, High Tempo", "Instrumentalness, Acousticness")
Song=c("Attention", "Caslte of the Glass", "Kun Faya Kun")
tbl2 <- data.frame(Clusters, Characteristics,Song)
library(kableExtra)
tbll2<-knitr::kable(tbl2, format = "html")
kable_styling(tbll2, bootstrap_options = c("striped", "hover","condensed"))
```

